---
title: "GapMinder"
author: "GG"
date: "3/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

```{r}
gapminder <- read_csv("https://raw.githubusercontent.com/jules32/2017-11-30-MBARI/gh-pages/data/gapminder.csv")
```

```{r}
head(gapminder)
tail(gapminder,2)
```


```{r}
str(gapminder)
```

```{r}
dim(gapminder)
ncol(gapminder)
nrow(gapminder)
```

```{r}
unique(gapminder$country)
unique(gapminder[, "continent"])
```


```{r}
ggplot(data=gapminder, aes(x=year, y=pop, color=continent)) + geom_point() 
```

```{r}
summary(gapminder)
```

```{r}
# install.packages("skimr")
library("skimr")
```

```{r}
skim(gapminder)
```

```{r}
filter(gapminder, country == "Switzerland")
```

##Dplyer
*filter and select*
```{r}
filter(gapminder, lifeExp < 30)
```

```{r}
filter(gapminder, country == "Rwanda", year == max(year))
```

```{r}
dplyr::select(gapminder, country, year, pop)
```

```{r}
dplyr::select(gapminder, -continent)
```

```{r}
gapminder_no_cont <- dplyr::select(gapminder, -continent)
filter(gapminder_no_cont, year == 2007)
```

##Easier way using pipe
*This notation signifies to use the result of the first statement in the second command.* 

```{r}
select(gapminder, -continent) %>% filter(year == 2007)
```

##Mutate
*A way to do an operation on an existing column and create a new version as a separate column*

```{r}
gapminder %>% mutate(gdp = gdpPercap * pop) %>% arrange(desc(gdp))
```

##Group by

*Allows you to use a subset of the dataframe*

```{r}
gapminder %>% group_by(country) %>% mutate(gdp = gdpPercap * pop, max_gdp = max(gdp)) %>% ungroup()
```

##Combining Group_By and Summarize

```{r}
gapminder %>% group_by(country) %>% mutate(gdp = gdpPercap * pop) %>%
  summarize(max_gdp = max(gdp)) %>% ungroup()
```

##Joining Datasets in Tidyr
*using left and right_join or full_join to combine tables. Order of datasets in the code assigns position, i.e. left or right. Use similar to merge in baser*

```{r}
co2 <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/co2.csv")
rm(Co2)
head(co2)
summary(co2)

```

*combine co2 dataset with gapminder*
```{r}
gap_2007 <- gapminder %>% filter(year == 2007) 
length(unique(gapminder$country))
left_join(gap_2007, co2, by = "country")

```

```{r}
right_join(gap_2007, co2, by = "country")
```

*can use set_diff() to see where the datasets don't match*

##Tidyr
*using Airpassengers dataset loaded with Tidyr*

```{r}
AirPassengers
```

*gather - will gather data from rows and combine into columns*
*spread - does the opposite - takes columns and spreads them into rows*
*separate - separates one column into several*
*unite - combines several columns into one*

```{r}
?gather
AirPassengers <- data.frame(AirPassengers)
gather(AirPassengers, month, passengers)
```


```{r}
gap_wide <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder_wide.csv")
```

*Let's try gather again*

```{r}
gap_wide %>% gather(observation, observ_value)
```

*too much collapse. Let's try again*

```{r}
gap_wide %>% gather(observation, observ_value, -continent, -country)
```

*Still better*
```{r}
gap_better <- gap_wide %>% gather(observation, observ_value, -continent, -country) %>% 
  separate(observation, into = c("obs_type", "year"), sep = "_", convert = TRUE)

unique(gap_better$obs_type)


```

*Even better*
```{r}
gap_best <- gap_better %>% spread(obs_type, observ_value)
gap_best
```








