---
title: "Tidy coral"
author: "GG"
date: "March 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## libraries
```{r}
library(tidyverse)
install.packages("janitor")
library(janitor) # install.packages('janitor')
library(skimr) # install.packages('skimr')
library(stringr) # added when we needed it for benthic data
```
## data filepaths/urls ----

## benthic data
benthic_url <- 'https://www.nodc.noaa.gov/archive/arc0054/0104255/1.1/data/0-data/cd08/100308OaAla03m.CSV'

## buoy data
buoy_url <- 'http://www.ndbc.noaa.gov/view_text_file.php?filename=mokh1h2010.txt.gz&dir=data/historical/stdmet/'

```{r}
benthic_raw <- read_csv(benthic_url)
head(benthic_raw)
```
##Clean up column names using the janitor package!

```{r}
benthic <- benthic_raw %>% janitor::clean_names()
names(benthic)

```

```{r}
head(benthic)
```
```{r}
benthic <- benthic %>% select(id_name, point, x, y, id_date)
head(benthic)

```
##get rid of the # signs
```{r}
benthic <- benthic %>% mutate(date = stringr::str_remove_all(id_date, "#"))
head(benthic)

```

```{r}
skimr::skim(benthic)
```
```{r}
ggplot(data = benthic, aes(x = date, fill = id_name)) + geom_bar()
```


##Buoy Data
first read in data
```{r}
buoy_raw <- read.csv(buoy_url)
head(buoy_raw)

```

```{r}
buoy_raw <- read_table2(buoy_url)
head(buoy_raw)

```

##remove the second header row

```{r}
buoy <- buoy_raw
## 1. overwrite column names
names(buoy) <- str_c(names(buoy),                  ## current header
                  buoy %>% filter(`#YY` == "#yr"), ## row1 -- don't say buoy[1,] 
                  sep = "_")                       ## separate by `_`

## inspect
names(buoy) ## Looks a lot better

```

```{r}
## 2. clean up a bit more to get rid of the `#`s and the `/`s. 
names(buoy) <- str_replace_all(names(buoy), "#", "")  # replace `#` with nothing 
names(buoy) <- str_replace_all(names(buoy), "/", "_") # replace `/` with `_`

## inspect to make sure it worked
names(buoy)
```

```{r}
## 3. remove redundant row with units
buoy <- buoy %>%
  filter(YY_yr != "#yr")

head(buoy)
```
 ##Now let's make a plot
 
```{r}
ggplot(buoy, aes(WTMP_degC)) + geom_bar() + theme(axis.text.x = element_text(angle = 90))
```
 
```{r}
buoy <- buoy %>% 
  unite(date, c(YY_yr, MM_mo, DD_dy), sep = "-")
head(buoy)
```

```{r}
bb_join <- left_join(benthic, buoy, by = "date")

bb_join %>% select(id_name, x, y, date, hh_hr, mm_mn, WTMP_degC) %>% head()

```

```{r}
rm(buoy_temp) 
```

```{r}
buoy_daily <- buoy %>%
  mutate(temp_clean = gsub("999.0", NA, WTMP_degC)) %>%
  mutate(temp_clean = as.numeric(temp_clean)) %>% #first need to get rid of 999's
  group_by(date) %>% 
  summarise(mean_temp = mean(temp_clean, na.rm=T)) #then get mean temperature
```

```{r}
head(buoy_daily)


```

```{r}
bb_join <- left_join(benthic, buoy_daily, by = "date")
bb_join
```
```{r}
ggplot(bb_join, aes(mean_temp)) + geom_histogram() 
```

